Key Provisioning for the Internet of Things via Light

### 2015-10-14 paper/paper.tex

The smartphone encodes the data string (e.g. ASCII encoding), measures its binary length and calculates the checksum.
All data that is being sent is encoded with a \textit{Hamming(8,4)-code} so that two-bit errors can be detected and one-bit errors can be corrected.
\textit{CRC32} is chosen as the checksum algorithm to avoid saving erroneous key material.
This could be caused by 3-or-more-bit errors that the Hamming code failed to detect.
The concatenated bit string is then sent using flashing of the display or the flash light.

@@

We have developed two different versions of the Android application, which enable users to enter keys and transmit them.
The first version is written using web technologies (i.e. HTML, CSS, and JavaScript) and deployed to the smartphone using Apache~Cordova\footnote{https://cordova.apache.org/}.
Because of this, the application can also run in most desktop browsers and can easily be ported to work with other smartphone operating systems, too, e.g. iOS or Windows Phone.
There are two disadvantages of this implementation, which motivated us to build a second edition:
The flash light cannot be controlled reliably using the mentioned web technologies.
This forced us to use the screen for light emission, which resulted in far smaller brightness margins to be read by the mote.
Furthermore the timing accuracy was considerably limited due to the overhead caused by the Cordova runtime and the JavaScript interpreter which resulted in transmission errors.

@@

To improve transmission reliability, we therefore built a native Android application written in Java.
Still, on Android, it is impossible to guarantee real-time behavior, but we get smaller delays without JavaScript involved.
As to bee seen in Fig.~\ref{fig:screenshot}, any entered text will be ASCII-encoded and transmitted.
This is just a prove of concept, a real key should be a random 128~bit string which needs to be generated and saved for later usage, which is out of scope of this paper.
This Java implementation benefits from the ability of the native Java timer to re-adjust itself, which means that timing inaccuracies do not add up.
In effect, the smartphone and the mote do not get out of sync, no matter how long the transmission takes.
Another advantage of this version is the flash light.
Depending on the smartphone model, the flash light is as fast as screen flickering even for small bit lengths, while being much brighter.
Especially when attempting to program multiple motes at once this comes in very handy.

@@

Once the light process is started, an LED-countdown (red $\rightarrow$ blue $\rightarrow$ green) indicates the start of the key initialization procedure.
After the successful calibration, synchronization and initialization a blue LED indicates that the mote is receiving data.
If an error is detected by the Hamming-code or if the computed checksum does not match the transmitted one, a red LED indidcates the error and the process should be restarted.
A green LED shows that the key material has been received successfully and was saved to the permanent flash memory of the mote.

\subsubsection{Contiki driver wrapper}
\label{ssub:contiki_driver_wrapper}

For the mote to actually use the key material that was transmitted, the initialization of the link layer security needs to be delayed until the key transmission finished successfully.
To do so we created a link layer driver wrapper that accomplishes the following things:

\begin{enumerate}
	\item Delay all calls to the link layer driver.
	\item Check whether the key material was initialized before.
	\item If not, start the \textit{light process} and wait for successful key transmission.
	\item Initialize the link layer using the transmitted key material.
	\item Forward all calls to the link layer driver.
\end{enumerate}

Calls to the driver that are being issued before the key material initialization are actually being canceled right now, but should eventually just be delayed and reissued once the initialization has finished.
Including the light app on a mote automatically redefines \textit{NETSTACK\_CONF\_LLSEC}\footnote{This global constant defines the link layer driver that is used by the mote.} to point to our link layer driver wrapper.
On the first start of the mote the key initialization via the \textit{light process} is triggered whereas during later starts the mote recognizes that its key material has been initialized before.

>>>>

The smartphone encodes the data string (e.g. ASCII encoding), measures its binary length and calculates the checksum.
All data that is being sent is encoded with a \textit{Hamming(8,4)-code} so that two-bit errors can be detected and one-bit errors can be corrected.
\textit{CRC32} is chosen as the checksum algorithm to avoid saving erroneous keying material.
This could be caused by 3-or-more-bit errors that the Hamming code failed to detect.
The concatenated bit string is then sent using flashing of the display or the flashlight.

@@

We have developed two different versions of the Android application, which enable users to enter keys and transmit them.
The first version is written using web technologies (i.e. HTML, CSS, and JavaScript) and deployed to the smartphone using Apache~Cordova\footnote{https://cordova.apache.org/}.
Because of this, the application can also run in most desktop browsers and can easily be ported to work with other smartphone operating systems, too, e.g. iOS or Windows Phone.
There are two disadvantages of this implementation, which motivated us to build a second edition:
The flashlight cannot be controlled reliably using the mentioned web technologies.
This forced us to use the screen for light emission, which resulted in far smaller brightness margins to be read by the mote.
Furthermore the timing accuracy was considerably limited due to the overhead caused by the Cordova runtime and the JavaScript interpreter which resulted in transmission errors.

@@

To improve transmission reliability, we therefore built a native Android application written in Java.
Still, on Android, it is impossible to guarantee real-time behavior, but we get smaller delays without JavaScript involved.
As to bee seen in Fig.~\ref{fig:screenshot}, any entered text will be ASCII-encoded and transmitted.
This is just a prove of concept, a real key should be a random 128~bit string which needs to be generated and saved for later usage, which is out of scope of this paper.
This Java implementation benefits from the ability of the native Java timer to re-adjust itself, which means that timing inaccuracies do not add up.
In effect, the smartphone and the mote do not get out of sync, no matter how long the transmission takes.
Another advantage of this version is the flashlight.
Depending on the smartphone model, the flashlight is as fast as screen flickering even for small bit lengths, while being much brighter.
Especially when attempting to program multiple motes at once this comes in very handy.

@@

Once the light process is started, an LED-countdown (red $\rightarrow$ blue $\rightarrow$ green) indicates the start of the key provisioning procedure.
After the successful calibration, synchronization and initialization a blue LED indicates that the mote is receiving data.
If an error is detected by the Hamming-code or if the computed checksum does not match the transmitted one, a red LED indidcates the error and the process should be restarted.
A green LED shows that the keying material has been received successfully and was saved to the permanent flash memory of the mote.

\subsubsection{Wrapper}
\label{ssub:contiki_driver_wrapper}

For the mote to actually use the keying material that was transmitted, the initialization of the link layer security needs to be delayed until the transmission finished successfully.
To do so we created a link layer driver wrapper that accomplishes the following things:

\begin{enumerate}
	\item Delay all calls to the link layer driver if the mote was not provisioned with keying material before.
	\item Start the \textit{light process} and wait for successful transmission.
	\item Initialize the link layer using the transmitted keying material.
	\item Forward all calls to the link layer driver.
\end{enumerate}

Including the light app on a mote automatically redefines \textit{NETSTACK\_CONF\_LLSEC}\footnote{This global constant defines the link layer driver that is used by the mote.} to point to our link layer driver wrapper.
At the first start of the mote, the keying material initialization via the \textit{light process} is triggered whereas during later starts the mote recognizes that its keying material has been initialized before.
